---
title: "Profit distributions in Iowa, 2012-2015"
author: "Elke Brandes"
date: "Tuesday, November 01, 2016"
output: html_document
---
This paper describes the calculation of profit distributions for 2012-2015 in Iowa.

Using dplyr, I can access the PostgreSQL datbase to import the tables I need to produce the histograms. 

1. Connect to the isuag database and load tables:
```{r}
library('dplyr')

# Connect to local PostgreSQL via dplyr
pw <- {
  "5QUXJHTbxj"
}
isuag_db <- src_postgres(dbname = 'isuag',
                           host = 'isu-ag-db.agsolver.com',
                           port = 5432,
                           user = 'isuag',
                           password = pw)

```
Use the dplyr PostgreSQL handle into the database:
```{r}

# this is not a data frame:
profit_mean_2012_2015 = tbl(isuag_db, "01_profit_mean_2012_2015_aggregated") 

# this is a data frame:
profit_mean_2012_2015_df = as.data.frame(profit_mean_2012_2015)
```
The `tbl()` command lets you access tables in the database remotely, and sql() lets you send queries. For now, I don't need to send queries to the database since I have done that already in the database. I only need to read in the tables. But I can run some tests here, for example to check how large the area is that is on average unprofitable:

```{r}
# send a query through dplyr
query = "SELECT SUM(sum_ha) FROM \"01_profit_mean_2012_2015_aggregated\" WHERE profit_mean_ha_rounded < 0"
sum = tbl(isuag_db, sql(query))
sum
```
Now I can load more tables from the database that I need for the annual profit distributions.
```{r}
file_list <- c("01_profit_rounded_aggregated_2012", "01_profit_rounded_aggregated_2013", "01_profit_rounded_aggregated_2014", "01_profit_rounded_aggregated_2015")


profit_2012 <- tbl(isuag_db,file_list[1]) %>%
as.data.frame()
profit_2013 <- tbl(isuag_db,file_list[2]) %>%
as.data.frame()
profit_2014 <- tbl(isuag_db,file_list[3]) %>%
as.data.frame()
profit_2015 <- tbl(isuag_db,file_list[4]) %>%
as.data.frame()

str(profit_2012)

```

After retrieving all tables I need, I shut the database down.
```{r}
rm(list=c('profit_mean_2012_2015','isuag_db')); gc() 
```

```{r}

data_list <- list(profit_2012, profit_2013, profit_2014, profit_2015)
str(data_list[[1]])
```
Packages needed for the plots:
```{r}
library("ggplot2")
library("gridGraphics") # needed for the function unit
```

I am using a modified theme for ggplot that has black border, white background, and inward facing tick marks:
```{r}
theme_b_border_inward <- function (base_size = 12, base_family = "") 
{
  theme_grey(base_size = base_size, base_family = base_family) %+replace% 
    theme(
      axis.text = element_text(size = rel(0.8), margin=unit(0.5, "cm")), 
      axis.ticks = element_line(colour = "black"),
          axis.ticks.length=unit(-0.15, "cm"),
#      axis.ticks.margin=unit(0.5, "cm"),
          legend.key = element_rect(colour = NA), panel.background = element_rect(fill = "white", 
           colour = NA), panel.border = element_rect(fill = NA, 
           colour = "black"), panel.grid.major = element_line(colour = NA, 
           size = 0.2), panel.grid.minor = element_line(colour = NA, 
           size = 0.5), strip.background = element_rect(fill = "grey80", 
           colour = "grey50", size = 0.2))
}
```

I prepare a data frame to be used for the facet plot:
```{r}
profit <- vector(mode="numeric", length=0)
 for (i in seq(data_list))
   profit <- append(profit, data_list[[i]][,1])
  
area <- vector(mode="numeric", length=0)
 for (i in seq(data_list))
   area <- append(area, data_list[[i]][,2])

   
profit_2012_2015 <- data.frame(year = c(rep("2012", sapply(data_list[1],nrow)), rep("2013", sapply(data_list[2],nrow)), 
                                        rep("2014", sapply(data_list[3],nrow)), rep("2015", sapply(data_list[4],nrow))),
                              profit,                             
                              area = area*1e-6) # this factor to show Mha

head(profit_2012_2015)
```

Then I plot the distributions:
```{r}
hi_profit <- ggplot(profit_2012_2015, aes(profit, weight = area)) + 
  geom_histogram(binwidth=50, alpha = .5, position="identity") +
  theme_b_border_inward() +
#  ylim(0,1500000) + xlim(-1500,1800) +
 scale_x_continuous(name="Profitability [US$/ha]") +
  scale_y_continuous(name="Area [Mha]") +
  theme(legend.position="bottom") +
  facet_wrap( ~ year, ncol=2) 

hi_profit +
#  scale_fill_manual(values=c("#E69F00", "#56B4E9"), name="", labels=c("Maize","Soybean")) +
  geom_vline(aes(xintercept= -250, linetype= "dashed"),  show.legend = FALSE) +
  geom_vline(aes(xintercept= 0, linetype= "dotted"),  show.legend = FALSE) 

```

