---
title: "Profit distributions in Iowa, 2012-2015"
author: "Elke Brandes"
date: "Tuesday, November 01, 2016"
output: html_document
---
This paper describes the calculation of profit distributions for 2012-2015 in Iowa.

Using dplyr, I can access the PostgreSQL datbase to import the tables I need to produce the histograms. 

1. Connect to the isuag database:
```{r}
library('dplyr')

# Connect to local PostgreSQL via dplyr
pw <- {
  "5QUXJHTbxj"
}
isuag_db <- src_postgres(dbname = 'isuag',
                           host = 'isu-ag-db.agsolver.com',
                           port = 5432,
                           user = 'isuag',
                           password = pw)


# this is not a data frame; it's a dplyr PostgreSQL handle into the database
profit_mean_2012_2015 = tbl(isuag_db, "01_profit_mean_2012_2015_aggregated") 

# this is a data frame
profit_mean_2012_2015_df = as.data.frame(profit_mean_2012_2015)
```
The tbl() command lets you access tables in the database remotely, and sql() lets you send queries. For now, I don't need to send queries to the database since I have done that already in the database. I only need to read in the tables. But I can run some tests here, for example to check how large the area is that is on average unprofitable:

```{r}
# send a query through dplyr
query = "SELECT SUM(sum_ha) FROM \"01_profit_mean_2012_2015_aggregated\" WHERE profit_mean_ha_rounded < 0"
sum = tbl(isuag_db, sql(query))

sum

# shut down database
rm(list=c('profit_mean_2012_2015','isuag_db')); gc() 
```

